---
title: 'Lesson 5: Reverse engineering your problem'
output:
  html_document:
    df_print: paged
objectives:
- ' To practice using the main dplyr verbs '
- ' To practice combining dplyr and ggplot'
- ' To understand how to go from desired outcome to solution'
---

```{r}
library(ggplot2)
library(tidyverse)
```


```{r}
gapminder <- read.csv("gapminder_data.csv", header=TRUE)
```

## Challenge 1

Using the gapminder data, recreate this figure

```{r echo=FALSE}
gapminder %>%
  filter(country == 'Nigeria' & year >= 1960 & year <= 2000) %>%
  ggplot(aes(year,gdpPercap)) +
  geom_line()
  

```

**Hints!**

gapminder %>%
  filter(_________) %>%
  ggplot(aes(______,________)) +
  geom_______
  
  

## Challenge 2

Using the gapminder data, recreate this table of the mean, min, and max life expectancy for each continent.

```{r echo=FALSE}
gapminder %>%
    group_by(continent) %>%
    summarize(
      meanLifeExp = mean(lifeExp),
      minLifeExp = min(lifeExp),
      maxLifeExp = max(lifeExp))
```

## Challenge 3

For the year 2000, recreate the following table showing:

- Content
- Mean GDP per capita
- Mean population 
- Mean GDP per billion people 

Hint: gdpPercap = GDP/population. First, compute GDP. Then divide this by 1 billion.


```{r echo=FALSE}
gapminder %>%
    filter(year == 2002) %>% 
    mutate(gdp_billion = gdpPercap*pop/10^9) %>%
    group_by(continent) %>%
    summarize(mean_gdpPercap = mean(gdpPercap),
              mean_pop = mean(pop),
              mean_gdp_billion = mean(gdp_billion))
```

## Challenge 4: Make this plot!

Make a plot of gdp per capita for Canada, Mexica, India, and the UK. 
Note some of the tricky details in the formatting...



```{r echo=FALSE}
gapminder %>%
   filter(country == 'Canada' | country == 'Mexico' | country == 'India' | country == 'United Kingdom') %>% 
   # Make the plot
   ggplot(aes(x = year, y = gdpPercap, color = continent)) +
   geom_line() +
   facet_wrap( ~ country,scales='free')
```


## Bonus challenge!!

Make this boxplot summarizing the meanLifeExpectancy for each continent (over all years)

```{r echo=FALSE}
gapminder %>%
   group_by(continent, year) %>%
   summarize(meanLifeExp = mean(lifeExp)) %>%
   # Make the plot
   ggplot(aes(x = continent, y = meanLifeExp, color = continent)) +
   geom_boxplot()
```

